---
title: "eds 221 day 4"
author: "Lewis White"
date: "2022-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```


## INTRO TO FUNCTIONS

```{r}
quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

qs_vec <- vector(mode = "numeric", length = length(quarter_splits))

for (i in seq_along(quarter_splits)) {
  consecuative_sum = quarter_splits[i] + quarter_splits[i + 1]
  qs_vec[i] <- consecuative_sum
}
```


```{r}
convert_animal_age <- function(animal, human_age) {
  if (animal == "dog") {
    print(human_age * 7)
  } else if (animal == "goat") {
    print(human_age * 4.7)
  } else {
    print("Pick again, sucker!")
  } 
}

convert_animal_age(animal = "cow", human_age = 10)
convert_animal_age(animal = "goat", human_age = 10)
```

make a function to double the user input

```{r}
double_this <- function(my_value) {
  print(2 * my_value)
}

double_this(my_value = 200)
double_this(40)

double_this(c(1, 3, 4, 9, 139))
```

your turn: write a function that asks for users name and age, and returns an exclamation like "I'm name and I am age years old

```{r}
intro <- function(name, age) {
  print(paste("I'm", name, "and I am", age, "years old!"))
}

intro(name = "Lana",age = 31)
```

make a function using a dataframe

```{r}
dog_choice <- tribble(
  ~dog_name, ~food,
  "Khora", "everything",
  "Teddy", "salmon",
  "Waffle", "pancakes",
  "Banjo", "chicken"
)

dog_menu <- function(name) {
  my_sub <- filter(.data = dog_choice, dog_name == name)
  print(paste("My name is", my_sub$dog_name, "and my favorite food is", my_sub[,2]))
}

dog_menu("Khora")



```


## functions meet for loops 

```{r}
means_over_columns <- function(df) {
  for (i in seq_along(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)
    column_name <- colnames(df[i])
    print(paste("The mean value of", column_name, "is", col_mean))
  }
}

means_over_columns(df = mtcars)
means_over_columns(df = faithful)

```


## Example nested for loop

```{r}
clothing <- c("Shirt", "pants", "shoes")
color <- c("purple", "red", "blue", "white")

our_clothes <- matrix(nrow = length(color), ncol = length(clothing))

for (i in seq_along(clothing)) {
  for (j in seq_along(color)) {
    clothing_details <- paste(color[j], clothing[i])
    our_clothes[j,i] <- clothing_details
  }
}
```


```{r}
sharks <- c("hammerhead", "great white", "thresher", "megladon")
snack <- c("tutle", "human", "golfish")

shark_snacks_df <- matrix(nrow = length(snack), ncol= length(sharks))

for (i in seq_along(sharks)) {
  for (j in seq_along(snack)) {
    sentence <- paste(sharks[i], "eat lots of", snack[j])
    shark_snacks_df[j, i] <- sentence
  }
}
```


## Logistic growth

Modeling a population over tume and varying growth rate 

```{r}
orange_rainbow_function <- function(oranges, rainbows) {
  (oranges * rainbows / 2.4)^2
}

#shortcut to create function =  command + option + x
```


```{r}
logistic_growth <- function(K, N0, r, time) {
  K / (1 + ((K - N0) / N0) * exp(-r * time))
}

#TEST
logistic_growth(K = 5000, N0 = 100, r = 0.07, time = 42)
```

```{r}
time_vec <- seq(from = 0, to = 400, by = 1)

pop_35 <- logistic_growth(N0 = 40, K = 8000, r = 0.05, time = time_vec)
pop_35

pop_time35_df <- data.frame(time_vec, pop_35)

ggplot(data = pop_time35_df, aes(x = time_vec, y = pop_35)) +
  geom_line()
```


```{r}
pop_35_vec <- vector(mode = "numeric", length = length(time_vec))

for (i in seq_along(time_vec)) {
  population <- logistic_growth(K = 6000, N0 = 300, r = 0.06, time = time_vec[i])
  pop_35_vec[i] <- population
}
 
```

## Now extend this to iterate over multiple growth rates

```{r}
r_seq <- seq(from = 0.01, to = 0.5, by = 0.05)
```

```{r}
out_matrix <- matrix(nrow = length(time_vec), ncol = length(r_seq))

for (i in seq_along(r_seq)) {
  for (j in seq_along(time_vec)) {
    pop <- logistic_growth(K = 6000, N0 = 300, r = r_seq[i], time = time_vec[j])
    out_matrix[j,i] <- pop
  }
}

### now some wrangling!!

logistic_df <- data.frame(time = time_vec, out_matrix) 

colnames(logistic_df) <- c(paste0("gr_", r_seq), "time")


logistic_df_long <- logistic_df %>%
  pivot_longer(cols = -time, values_to = "population", names_to = "growth_rate")
  
```


```{r}
ggplot(data = logistic_df_long, aes(x = time, y = population)) + 
  geom_line(aes(color = growth_rate))
```

Use `stop` to add an error



$$
P = 0.3\rho A v^3
$$

Rho: air density (kg / m^3)
A: area covered by turbine (square meters)
V: wind velocity (m/s)

```{r}
calc_windpower <- function(rho, radius, windspeed) {
  
  if (windspeed > 130) {
    warning("Whoa, thant's really fast ~ are you sure?")
  }
  
  if (rho > 1.225) {
    warning("That air density is sus")
  }
  
  if (radius < 0) {
    stop("radius must be positive")
  }
  
  0.3 * rho * pi * (radius^2) * (windspeed^3)
}

calc_windpower(rho = 2.225, radius = 1, 20)
```


